---
title: "Hw2"
author: "Yuan Meng"
date: 2021-10-7
output: github_document
---

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1

```{r message = FALSE}
trash_data = read_excel("/Users/yuanmeng/Desktop/Fall 2021/P8105 Data Science/DS_Hw/p8105_hw2_ym2866/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", range = "A2:N535") #Read Excel file and omit non-data entries
trash_data
```

```{r}
trash_data= janitor::clean_names(trash_data) 
names(trash_data) #use reasonable variable names
drop_na(trash_data, dumpster) #omit rows that do not include dumpster-specific data
mutate(trash_data, sports_balls = round(sports_balls)) #round the number of sports balls to the nearest integer
```

```{r message = FALSE}
precipitation_19 = read_excel("/Users/yuanmeng/Desktop/Fall 2021/P8105 Data Science/DS_Hw/p8105_hw2_ym2866/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 6, range = "A2:B14") #omit rows without precipitation data
precipitation_18 = read_excel("/Users/yuanmeng/Desktop/Fall 2021/P8105 Data Science/DS_Hw/p8105_hw2_ym2866/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 7, range = "A2:B14") #omit rows without precipitation data
precipitation_19 = mutate(precipitation_19, year = "2019") #add a variable for year
precipitation_18 = mutate(precipitation_18, year = "2018") #add a variable for year
precipitation_19
precipitation_18
```

```{r}
prec_data = bind_rows(precipitation_19, precipitation_18) # combine precipitation datasets
prec_data
as.character(month.name) #convert month to a character variable
```
The number of observation in Trash Wheel Collection Totals data is on 453 dumpster.The data include `r ncol(trash_data)-1` key variables, they are  "month", "year", "date", "weight_tons", "volume_cubic_yards", "plastic_bottles", "polystyrene", "cigarette_butts", "glass_bottles", "grocery_bags", "chip_bags", "sports_balls", "homes_powered".  

The number of total observation in combine 2019 and 2018 precipitation datasets is `r nrow(prec_data)`. The data include `r ncol(prec_data)` variables. They are month total and year. The total precipitation in 2018 is 70.33 and total  precipitation in 2019 is 33.95.  The median number of sports balls in a dumpster in 2019 is `r median(as.numeric(unlist(precipitation_19[1:12,2])))`.

# Problem 2
```{r message = FALSE}
pols_month = read_csv("./fivethirtyeight_datasets/pols-month.csv")
pols_month = separate(pols_month,1,c("year","month","day"),sep="-") # break up the variable into integer variables
pols_month = pols_month %>% 
  mutate(month = month.name[as.numeric(month)]) #replace month number with month name
pols_month = mutate(pols_month, president = ifelse(prez_gop == 1, "gop","dem"))
pols_month = select(pols_month, -c(prez_dem, prez_gop, day)) #remove prez_dem and prez_gop; and remove the day variable.
pols_month
```

```{r}
snp_data = read_csv("./fivethirtyeight_datasets/snp.csv")
snp_data = separate(snp_data,1,c("month","day","year"),sep="/") #break up the variable into integer variables year, month, and day;
snp_data = snp_data %>% 
  mutate(month = month.name[as.numeric(month)]) #replace month number with month name
snp_data = snp_data [, c("year", "month", "day","close")] #organize let year and month are the leading columns.
snp_data
```

```{r message = FALSE}
unemployment_data = read_csv("./fivethirtyeight_datasets/unemployment.csv")
unemployment_data = unemployment_data %>%
  pivot_longer(!Year, names_to = "Month", values_to = "unemployment_rate")#switching from “wide” to “long” format
unemployment_data
```

```{r}
join_pols_snp = full_join(pols_month,snp_data )#merging snp into pols
join_unemployment = bind_rows(join_pols_snp,unemployment_data) #merging unemployment into the result
join_pols_snp
join_unemployment
```
The pols_month dataset have 11 variables and 1609 observations.The key variables are year, month,president,gov_gop,sen_gop,rep_gop,gov_dem,sen_dem,rep_dem. The snp dataset have 4 variables and 787 observations. The key variables are year,month,day,close. The unemployment dataset have 3 variables and 816 observations. The key variables are year month and unemployment rate. The dataset merge snp, pols and unemployment have 14 variables and 533 observation. The key variables are variables from pols_month, snp_data and unemployment_data.

# Problem3
```{r}
popular_Baby_Names = read.csv("./Popular_Baby_Names.csv")
popular_Baby_Names = 
  popular_Baby_Names %>% 
  mutate(Child.s.First.Name = tolower(Child.s.First.Name),Gender = tolower(Gender),Ethnicity = tolower(Ethnicity )) #change children's name gender and ethnicity to lower case letter
  popular_Baby_Names["Ethnicity"][popular_Baby_Names["Ethnicity"]== "asian and paci"] <-"asian and pacific islander"
  popular_Baby_Names["Ethnicity"][popular_Baby_Names["Ethnicity"]== "black non hisp"] <-"black non hispanic"
  popular_Baby_Names["Ethnicity"][popular_Baby_Names["Ethnicity"]== "white non hisp"] <-"white non hispanic" # Under ethnicity there are lots of inconsistent ethnicity names. Change then to consistent 
popular_Baby_Names = popular_Baby_Names %>% distinct() #delete duplicated rows
popular_Baby_Names
```

```{r}
olivia_data = filter(popular_Baby_Names, Child.s.First.Name == "olivia") #filter out children who name Olivia
olivia_data = select(olivia_data, -c(Child.s.First.Name, Count, Gender)) #delete gender, count,name, columns 
olivia_data = olivia_data %>%
  pivot_wider(names_from = Year.of.Birth, values_from = Rank) # let the dataset have rows for ethnicity and columns for year
olivia_data
```


```{r}
male_data = filter(popular_Baby_Names, Gender == "male" & Rank == 1 ) #filter out male children
male_data = select(male_data, -c(Count, Gender, Rank))  #delete gender, count,name, columns
male_data = male_data %>% pivot_wider(names_from = Year.of.Birth, values_from = Child.s.First.Name)# let the dataset have rows for ethnicity and columns for year
male_data
```


```{r}
names_plot = filter(popular_Baby_Names, Gender == "male" & Ethnicity == "white non hispanic" & Year.of.Birth == 2016) #filter male, white non-hispanic children born in 2016
ggplot(names_plot, aes(x = Rank , y = Count)) + geom_point() + geom_text(aes(label = Child.s.First.Name), alpha = 0.5)# scatter plot showing the number of children with a name against the rank in popularity of that name
labs(title = "number of white non hispanic male children's name rank in 2016")
ggsave("8105_hw2_ym2866.pdf")
```



